Probability

### 2016-04-25 prob.tex
projections

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases. \\
\yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
using base 16 directly.}\\
Therefore, in a resource-aware setting, it is
more appropriate to consider probability measures that map events to a
finite set of elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the half-open interval
$(0,1]$, and $\vee$ as the addition of intervals. The definition of a
probability measure in this case is modified as being a function
$\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

To properly explain the previous example and generalize to arbitrary
quantum systems, we formally define a quantum probability space as
follows.

\begin{definition}[Quantum Probability Space~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}]
  Given a Hilbert space $\Hilb$, a \emph{quantum probability space}
  consists of a set of events $\events$ and a probability measure
  $\mu : \events \rightarrow [0,1]$ such that:\footnote {It is
    possible to define a more general space of events consisting of
    all operators~$\mathcal{A}$ on $\Hilb$ and consider
    $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.  When
    an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$ is
    the expectation value of $A$. We does not take this approach
    because we want to focus only on probability. }
\begin{itemize}
\item The set of events consists of the empty projection $\mathbb{0}$
  together with projection operators $\proj{\psi}$ for each state
  $\ket{\psi}$,\\
  \yutsung{and the sum of the orthogonal events recursively.}
\item $\mu(\mathbb{1})=1$, and 
\item for projections expressible as sums $\sum_{i}E_{i}$ of
  necessarily orthogonal events $E_i$, we have
  $\mu\left(\sum_{i}E_{i}\right)=\sum_{i}\mu\left(E_{i}\right)$.
\end{itemize}
\qed\end{definition}

>>>>

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware setting, it is more appropriate to
consider probability measures that map events to a finite set of
elements computable with a fixed set of
resources~\cite{Artstein1972,PuriRalescu1983}. The simplest such set,
and the one we will consider exclusively in this paper, is the set
$\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with the
operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the first definition, one can interpret
\imposs\ as the closed interval $[0,0]$, \poss\ as the half-open
interval $(0,1]$, and $\vee$ as the addition of intervals. The
definition of a probability measure in this case is modified as being
a function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

@@

To properly explain the previous example and generalize to arbitrary
quantum systems, we formally discuss projection operators and then
define a quantum probability space.

\amr{

  Projections include $\mathbb{0}, \proj{\psi}$ and $P_1 + P_2$ and
  $P_1 . P_2$ if $P_1$ and $P_2$ commute.

  When we write an event as $\sum E_i$ we are implicitly assuming they
  are commutative.

  Orthogonal implies commutative but not vice-versa
}

\begin{definition}[Quantum Probability Space~\cite{BirkhoffVonNeumann1936,gleason1957,Redhead1987-REDINA,DBLP:journals/corr/abs-0910-2393,Maassen2010}]
  Given a Hilbert space $\Hilb$, a \emph{quantum probability space}
  consists of a set of events $\events$ and a probability measure
  $\mu : \events \rightarrow [0,1]$ such that:\footnote {It is
    possible to define a more general space of events consisting of
    all operators~$\mathcal{A}$ on $\Hilb$ and consider
    $\mu:\mathcal{A}\rightarrow\C$~\cite{Maassen2010,Swart2013}.  When
    an operator $A\in\mathcal{A}$ is Hermitian, $\mu\left(A\right)$ is
    the expectation value of $A$. We does not take this approach
    because we want to focus only on probability. }
\begin{itemize}
\item The set of events consists of all projections. This set includes
  the empty projection, projection operators $\proj{\psi}$ for each
  state $\ket{\psi}$, and sums of \emph{commuting} projections, 
\item $\mu(\mathbb{1})=1$, and 
\item for mutually orthogonal projections $E_i$, we have
  $\mu\left(\sum_{i}E_{i}\right)=\sum_{i}\mu\left(E_{i}\right)$.
\end{itemize}
\qed\end{definition}

%% See http://math.stackexchange.com/questions/1146893/on-the-sum-of-projection-operators
% Two properties: E1 orthogonal to E2 or not
%                          E1 commutative with E2 or not

% the empty projection $\mathbb{0}$

%   \yutsung{and the sum of the orthogonal events recursively.}
